---
title: "Data Import"
output: github_document
---


```{r}
library(tidyverse)
library(readxl)
library(haven)
```

## Read in some data

Read in the litters dataset.

```{r}
litters_df = read_csv ("./data/FAS_litters.csv")
litters_df = janitor :: clean_names (litters_df)
```

## Take a look at the data

Printing in the console

```{r}
litters_df
```

Show the first 6 rows or the last 6 rows
```{r}
head (litters_df)
tail (litters_df)
```

```{r}
skimr:: skim (litters_df)
```
Could also use view function to look at the dataset in the console.

## Options to read csv

Check out `?read.csv()` for more information

```{r}
litters_df= read_csv ("./data/FAS_litters.csv", skip = 10, col_names = FALSE)
litters_df= read_csv ("./data/FAS_litters.csv", na = c("", ".", "NA", 999))
```

## Other file formats

# Excel

Read in excel file
Look at `?read_excel()` for more information

```{r}
mlb_df = read_excel ("./data/mlb11.xlsx", range = "A1:F7")
```

# SAS

Read in SAS file

```{r}
pulse_df = read_sas ("./data/public_pulse_data.sas7bdat")
```

## Comparison with Base R

What about `read.csv` ?
Don't use this ever. Always use the _ notation

```{r}
litters_base = read.csv ("./data/FAS_litters.csv")
litters_readr = read_csv("./data/FAS_litters.csv")

litters_base
litters_readr
```

## Exporting data

Export the mlb sub-table

```{r}
write_csv (mlb_df, "./data/mlb_subtable.csv")
```

